Breast Cancer Diagnosis Classification Analysis

### 2022-12-08 austins_temp_powerpoint.rmd
updated

```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
train_control <- trainControl(method = "cv", number = 10, preProcOptions = list(thresh=0.99), classProbs=TRUE, summaryFunction = twoClassSummary)
```

@@

```{r}
acc_test <- numeric() 
for(i in 1:30){
    predict <- knn(train=training[,-1], test=test[,-1], cl=training$diagnosis, k=i, prob=T)
    acc_test <- c(acc_test,mean(predict==test[,1]))
}
```

>>>>

```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
```

@@

#TUNNING MODELS

# KNN

## Tuning KNN

```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
type_two_error <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)

for (i in 1:nrow(type_two_error)) {
    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error$k[i]))
    knn_pred <- predict(knn, test)
    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
    type_two_error$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
    type_two_error$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
}
print(which.min(type_two_error$typetwoerror))
```

```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() + geom_line(aes(x=type_two_error$k, y=type_two_error$error), size=1) + geom_line(aes(x=type_two_error$k, y=type_two_error$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
```

## Tuned KNN

```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error$typetwoerror)))
knn_pred <- predict(tuned_knn, test)
knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
```

```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(knn_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "KNN Confusion Matrix", x = "Reference", y = "Prediction") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=2+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

# Random Forrest

## Tuning Random Forrest

```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
type_two_error <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)

for (i in 1:nrow(type_two_error)) {
    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error$ntree[i])
    rf_pred <- predict(rf, test)
    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
    type_two_error$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
    type_two_error$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
}
```

```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() + geom_line(aes(x=type_two_error$ntree, y=type_two_error$error), size=1) + geom_line(aes(x=type_two_error$ntree, y=type_two_error$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
```

## Tuned Random Forrest

```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = which.min(type_two_error$typetwoerror))
rf_pred <- predict(tuned_rf, test)
rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
```

```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(rf_confusion$table), aes(x = Reference, y = rev(Prediction), fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Random Forrest Confusion Matrix", x = "Reference", y = "Prediction") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=2+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

# SVM

## Tuning SVM

```{r svm_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
gamma <- seq(0.01, 100, length.out=7)
cost <- seq(0.01, 100, length.out=7)
errorLog <- data.frame(gamma = gamma, cost = cost, error = NA, typetwoerror = NA)

for (i in 1:length(gamma)) {
  for (j in 1:length(cost)) {
    svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric = "Accuracy", gamma = gamma[i], cost = cost[j])
    svm_pred <- predict(svm, test)
    svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive = "M")
    tempGamma <- gamma[i]
    tempCost <- cost[j]
    tempError <- (svm_confusion$table[1, 2] + svm_confusion$table[2, 1])/sum(svm_confusion$table)
    tempTwo <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
    errorLog <- rbind(errorLog, data.frame(gamma = tempGamma, cost = tempCost, error = tempError, typetwoerror = tempTwo))
    print(paste("completed", i, j))
  }
}
errorLog <- tail(errorLog, -7)
```

```{r svm_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(errorLog, aes(x=gamma, y=cost, fill= typetwoerror)) + 
  scale_fill_gradient(low="white", high="blue") +
  geom_tile()
```

## Tuned SVM

```{r tuned_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
svm2 <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric = "Accuracy", gamma = 0.01, cost = 0.01)
svm_pred <- predict(svm2, test)
svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive = "M")
print(svm_confusion)
```

```{r tuned_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}
```
